---
title: Task 2 bait-boat catch and effort data atlas for the AOTTP (bigeye, skipjack
  and yellowfin)
author: "Doug Beare"
date: "11 December 2015"
output: pdf_document
header-includes: \usepackage{tabularx}
force_captions: yes
subtitle: null
Grmd::docx_document:
  fig_caption: yes
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=10, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```


```{r attach libraries and connect to DB, include =F, echo=F,eval=T, message=FALSE,error=FALSE,warning=FALSE}
library(spatial)
library(sp)
library(doBy)
library(rgdal)
library(RODBC)
library(RColorBrewer)
library(ggplot2)
library(vmstools)
library(mgcv)
library(maps)
library(mapdata)
library(reshape2)
library(rgeos)
library(lattice)
library(pander)
library(kfigr)
library(effdisR)
chan <- odbcConnect("effdis-tuna-cc1", case="postgresql", believeNRows=FALSE)
```



```{r get bb data by flag,include=TRUE,echo=F,eval=T}
effort_type_by_flag_bb <- sqlQuery(chan, "SELECT flagname AS Flag, 
eff1type AS Effort_type, count(eff1type) as No_records
                 FROM t2ce_new
                 WHERE region ='AT' AND geargrpcode = 'BB' AND yearc = 2014
                 GROUP BY flagname, eff1type
                 ORDER BY flagname, eff1type, No_records;")
```


```{r get bait-boat data,include=TRUE,echo=F,eval=T}
bb <- sqlQuery(chan,"SELECT yearc AS year, trend, timeperiodid AS month, 
flagname, region, geargrpcode,lon,lat,longitude,latitude,quadid AS quad, squaretypecode AS square, catchunit, dsettype, eff1, eff1type
FROM t2ce_new
WHERE region ='AT' AND yearc > 2009 AND timeperiodid < 13 
AND geargrpcode = 'BB';")
df <- data.frame(quad=bb$quad,lat=bb$lat,lon=bb$lon,square=bb$square)
df$square <- as.character(df$square)
ldf <- length(df$square)
df1<- data.frame(quad=rep(NA,ldf),lat=rep(NA,ldf),lon=rep(NA,ldf),square=rep(NA,ldf))
for(i in 1:length(df[,1]))
{df1[i,] <- latLon(x=df[i,])}
bb$longitude<- df1$lon
bb$latitude <- df1$lat

```


```{r extract kg bigeye caught by selected flags from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE,eval=T}
bet <- sqlQuery(chan,"SELECT yearc AS year, trend,quadid as quad, squaretypecode AS square,
timeperiodid AS month, flagname, region, geargrpcode,lon,lat,longitude,latitude, 
catchunit, dsettype, eff1, eff1type, bet as measured_catch
FROM t2ce_new
WHERE region ='AT' AND timeperiodid < 13 AND yearc > 2009 AND 
geargrpcode IN ('LL','PS','BB') AND catchunit != '--' ;")
bet$species <- 'bet'
df <- data.frame(quad=bet$quad,lat=bet$lat,lon=bet$lon,square=bet$square)
df$square <- as.character(df$square)
ldf <- length(df$square)
df1<- data.frame(quad=rep(NA,ldf),lat=rep(NA,ldf),lon=rep(NA,ldf),square=rep(NA,ldf))
for(i in 1:ldf){df1[i,] <- latLon(x=df[i,])}
bet$longitude<- df1$lon
bet$latitude <- df1$lat

```



```{r extract kg skipjack caught by selected flags from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE,eval=T}
skj<- sqlQuery(chan,"SELECT yearc AS year, trend,quadid as quad, squaretypecode AS square,
timeperiodid AS month, flagname, region, geargrpcode,lon,lat,longitude,latitude, 
catchunit, dsettype, eff1, eff1type, skj as measured_catch
FROM t2ce_new
WHERE region ='AT' AND timeperiodid < 13 AND yearc > 2009 AND 
geargrpcode IN ('LL','PS','BB') AND catchunit != '--' ;")
skj$species <- 'skj'

df <- data.frame(quad=skj$quad,lat=skj$lat,lon=skj$lon,square=skj$square)
df$square <- as.character(df$square)
ldf <- length(df$square)
df1<- data.frame(quad=rep(NA,ldf),lat=rep(NA,ldf),lon=rep(NA,ldf),square=rep(NA,ldf))
for(i in 1:ldf){df1[i,] <- latLon(x=df[i,])}
skj$longitude<- df1$lon
skj$latitude <- df1$lat

```


```{r extract kg yellowfin caught by selected flags from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE,eval=T}
yft<- sqlQuery(chan,"SELECT yearc AS year, trend,quadid as quad, squaretypecode AS square,
timeperiodid AS month, flagname, region, geargrpcode,lon,lat,longitude,latitude, 
catchunit, dsettype, eff1, eff1type, yft as measured_catch
FROM t2ce_new
WHERE region ='AT' AND timeperiodid < 13 AND yearc > 2009 AND 
geargrpcode IN ('LL','PS','BB') AND catchunit != '--' ;")
yft$species <- 'yft'
df <- data.frame(quad=yft$quad,lat=yft$lat,lon=yft$lon,square=yft$square)
df$square <- as.character(df$square)
ldf <- length(df$square)
df1<- data.frame(quad=rep(NA,ldf),lat=rep(NA,ldf),lon=rep(NA,ldf),square=rep(NA,ldf))
for(i in 1:ldf){df1[i,] <- latLon(x=df[i,])}
yft$longitude<- df1$lon
yft$latitude <- df1$lat

```


#Task 2 bait-boat data for The Atlantic


*Table 1. Effort type sampling by flag for purse-seiners in EFFDIS Task II database in 2014*
```{r make table for bb data by flag,include=TRUE,echo=FALSE}
pander(effort_type_by_flag_bb)
```

##Distribution of Task 2 bait-boat data


```{r plot distribution of spanish bait-boat data,include=TRUE,echo=F}
par(mfrow=c(2,2),mar=c(1,1,2,1),oma=c(3,3,3,3))
spatial.coverage.by.year.task2(tdata=bb,start.year=2011,end.year=2014,
                                 which.gear='BB',which.flag='EU.Espa単a')
```

**Figure 1.** *Distribution of data: Spanish registered bait-boats between 2011 and 2014*


```{r plot distribution of Brazil bait-boat data,include=TRUE,echo=F}
par(mfrow=c(2,2),mar=c(1,1,2,1),oma=c(3,3,3,3))
spatial.coverage.by.year.task2(tdata=bb
                               ,start.year=2011,end.year=2014,
                                 which.gear='BB',which.flag='Brazil')
```
**Figure 2.** *Distribution of data: Brazilian bait-boats between 2011 and 2014*


```{r plot distribution of french baitboat data,include=TRUE,echo=F}
par(mfrow=c(2,2),mar=c(1,1,2,1),oma=c(3,3,3,3))
spatial.coverage.by.year.task2(tdata=bb,start.year=2011,end.year=2014,
                                 which.gear='BB',which.flag="EU.France")
```

**Figure 3.** *Distribution of data: French bait-boats between 2011 and 2014*

```{r plot distribution of portuguese baitboat data,include=TRUE,echo=F}
par(mfrow=c(2,2),mar=c(1,1,2,1),oma=c(3,3,3,3))
spatial.coverage.by.year.task2(tdata=bb,start.year=2011,end.year=2014,
                                 which.gear='BB',which.flag="EU.Portugal")
```

**Figure 4.** *Distribution of data: Portuguese bait-boats between 2011 and 2014*


```{r plot distribution of senegalese bait-boat data,include=TRUE,echo=F}
par(mfrow=c(2,2),mar=c(1,1,2,1),oma=c(3,3,3,3))
spatial.coverage.by.year.task2(tdata=bb,start.year=2011,end.year=2014,
                                 which.gear='BB',which.flag="Senegal")

```

**Figure 5.** *Distribution of data: Senegalese bait-boats between 2011 and 2014*


```{r plot distribution of south african bait-boat data,include=TRUE,echo=F}
par(mfrow=c(2,2),mar=c(2,2,2,2),oma=c(3,3,3,3))
spatial.coverage.by.year.task2(tdata=bb,start.year=2011,end.year=2014,
                                 which.gear='BB',which.flag="South Africa")
```

**Figure 6.** *Distribution of data: South African bait-boats between 2011 and 2014*


```{r plot distribution of ghanaian bait-boats data,include=TRUE,echo=F}
par(mfrow=c(2,2),mar=c(2,2,2,2),oma=c(3,3,3,3))
spatial.coverage.by.year.task2(tdata=bb,start.year=2011,end.year=2014,
                                 which.gear='BB',which.flag="Ghana")
```

**Figure 7.** *Distribution of data: Ghanaian bait-boats between 2011 and 2014*


```{r plot distribution of Venezuelan bait-boat data,include=TRUE,echo=F}
par(mfrow=c(2,2),mar=c(2,2,2,2),oma=c(3,3,3,3))
spatial.coverage.by.year.task2(tdata=bb,start.year=2011,end.year=2014,
                                 which.gear='BB',which.flag="Venezuela")
```

**Figure 8.** *Distribution of data: Venezuelan bait-boats between 2011 and 2014*


##Bigeye caught by bait-boat (2014).


```{r weight of bigeye caught by location in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,3),mar=c(2,2,2,2))
three.d.catch.by.year(tdata=bet,which.year=2014,which.gear='BB',gridx=1,gridy=1,which.species='bet',which.flag='EU.France',scaling.f=10,catchunit='kg')  
three.d.catch.by.year(tdata=bet,which.year=2014,
                        which.gear='BB',which.species='bet',gridx=1,gridy=1,which.flag='EU.Espa単a',scaling.f=10,catchunit='kg')
three.d.catch.by.year(tdata=bet,which.year=2014,
                        which.gear='BB',which.species='bet',gridx=1,gridy=1,which.flag='Venezuela',scaling.f=10,catchunit='kg')  
three.d.catch.by.year(tdata=bet,which.year=2014,which.gear='BB',which.species='bet',gridx=1,gridy=1,which.flag='South Africa',scaling.f=10,catchunit='kg')

three.d.catch.by.year(tdata=bet,which.year=2014,which.gear='BB',which.species='bet',gridx=1,gridy=1,which.flag='Senegal',scaling.f=10,catchunit='kg')

```

**Figure 9.** *Bigeye caught by selected nations by bait-boat in 2014*


##Skipjack caught by bait-boat in 2014.



```{r skipjack caught by location in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,3),mar=c(2,2,2,2))
three.d.catch.by.year(tdata=skj,which.year=2014,which.gear='BB',gridx=1,gridy=1,which.species='skj',which.flag='EU.France',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='BB',which.species='skj',gridx=1,gridy=1,which.flag='EU.Espa単a',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='BB',which.species='skj',gridx=1,gridy=1,which.flag='Senegal',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=skj,which.year=2014,which.gear='BB',which.species='skj',gridx=1,gridy=1,which.flag='South Africa',scaling.f=10,catchunit='kg')

three.d.catch.by.year(tdata=skj,which.year=2014,which.gear='BB',which.species='skj',gridx=1,gridy=1,which.flag='Venezuela',scaling.f=10,catchunit='kg')

three.d.catch.by.year(tdata=skj,which.year=2014,which.gear='BB',which.species='skj',gridx=1,gridy=1,which.flag='Brazil',scaling.f=10,catchunit='kg')

```

**Figure 10.** *Skipjack caught by bait-boat by 6 nations in 2014*


##Yellowfin caught by bait-boat in 2014.

```{r yellowfin caught by baitboat by location in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,3),mar=c(2,2,2,2))
three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='BB',gridx=1,gridy=1,which.species='yft',which.flag='EU.France',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='BB',which.species='yft',gridx=1,gridy=1,which.flag='EU.Espa単a',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='BB',which.species='yft',gridx=1,gridy=1,which.flag='Senegal',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='BB',which.species='yft',gridx=1,gridy=1,which.flag='South Africa',scaling.f=100,catchunit='kg')

three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='BB',which.species='yft',gridx=1,gridy=1,which.flag='Venezuela',scaling.f=100,catchunit='kg')

three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='BB',which.species='yft',gridx=1,gridy=1,which.flag='Brazil',scaling.f=100,catchunit='kg')


```

**Figure 11.** *Yellowfin tuna caught by bait-boat by 6 nations*








