---
title: Task 2 effort and catch data atlas for the AOTTP (bigeye, skipjack and yellowfin
  caught by long-line, purse-seine and bait-boat)
author: "Doug Beare"
date: "9 December 2015"
output: pdf_document
header-includes: \usepackage{tabularx}
subtitle: null
Grmd::docx_document:
  fig_caption: yes
  force_captions: yes
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=10, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

#Background 

The overall objective of the Atlantic Tuna Tagging Programme (AOTTP) is to contribute to food security and economic growth of the Atlantic developing coastal states by ensuring sustainable management of tropical tuna resources in the Atlantic Ocean
The specific objective of this programme is to provide evidence based scientific advices to developing costal states, and other Contracting Parties, to support the adoption of effective Conservation and Management Measures (CMMs) in the framework of the International Commission for the Conservations of Atlantic Tunas (ICCAT). This will be achieved through improving the estimation, derived from tag-recapture data, of key parameters for stock assessment analyses, i.e. growth, natural mortality, movements and stock structure, etc.



```{r attach libraries and connect to DB, include =F, echo=F,eval=T, message=FALSE,error=FALSE,warning=FALSE}
library(spatial)
library(sp)
library(doBy)
library(rgdal)
library(RODBC)
library(RColorBrewer)
library(ggplot2)
library(vmstools)
library(mgcv)
library(maps)
library(mapdata)
library(reshape2)
library(rgeos)
library(lattice)
library(pander)
library(kfigr)
library(effdisR)
chan <- odbcConnect("effdis-local", case="postgresql", believeNRows=FALSE)
```


```{r get ll data by flag,include=TRUE,echo=F, eval=T}
effort_type_by_flag_ll <- sqlQuery(chan, "SELECT flagname AS Flag, 
eff1type AS Effort_type, count(eff1type) as No_records
                 FROM t2ce
                 WHERE region ='AT' AND geargrpcode = 'LL' AND yearc = 2014
                 GROUP BY flagname, eff1type
                 ORDER BY flagname, eff1type, No_records;")
```


```{r get ps data by flag,include=TRUE,echo=F,eval=T}
effort_type_by_flag_ps <- sqlQuery(chan, "SELECT flagname AS Flag, 
eff1type AS Effort_type, count(eff1type) as No_records
                 FROM t2ce
                 WHERE region ='AT' AND geargrpcode = 'PS' AND yearc = 2014
                 GROUP BY flagname, eff1type
                 ORDER BY flagname, eff1type, No_records;")
```



```{r get bb data by flag,include=TRUE,echo=F,eval=T}
effort_type_by_flag_bb <- sqlQuery(chan, "SELECT flagname AS Flag, 
eff1type AS Effort_type, count(eff1type) as No_records
                 FROM t2ce
                 WHERE region ='AT' AND geargrpcode = 'BB' AND yearc = 2014
                 GROUP BY flagname, eff1type
                 ORDER BY flagname, eff1type, No_records;")
```


```{r get ll data for ten most important flags,include=TRUE,echo=F,eval=T}
ll <- sqlQuery(chan,"SELECT yearc AS year, trend, timeperiodid AS month, 
flagname, region, geargrpcode,lon,lat,longitude,latitude,quadid AS quad, squaretypecode AS square, catchunit, dsettype, eff1, eff1type
FROM t2ce
WHERE region ='AT' AND yearc > 2009 AND timeperiodid < 13 AND eff1type='NO.HOOKS' AND geargrpcode = 'LL' 
AND flagname IN ('Japan','Chinese Taipei','Brazil','Angola','EU.España','EU.Portugal','St. Vincent and Grenadines','Vanuatu','U.S.A.','Namibia') 
AND catchunit != '--' ;")


```


```{r get purse-seine data for select flags,include=TRUE,echo=F,eval=T}
ps <- sqlQuery(chan,"SELECT yearc AS year, trend, timeperiodid AS month, 
flagname, region, geargrpcode,lon,lat, longitude,latitude,quadid AS quad, squaretypecode AS square, catchunit, dsettype, eff1, eff1type
FROM t2ce
WHERE region ='AT' AND yearc > 2009 AND timeperiodid < 13 AND eff1type IN ('D.FISH','FISH.HOUR') 
AND geargrpcode = 'PS' AND flagname IN ('Belize','Côte D''Ivoire','EU.España','Guatemala','Guinée Rep.','Panama','Venezuela','Panama','EU.France','EU.Portugal','U.S.A.', 'Curaçao','EU.Netherlands','Maroc','Namibia' ) 
AND catchunit != '--' ;")




```



```{r get bait-boat data,include=TRUE,echo=F,eval=T}
bb <- sqlQuery(chan,"SELECT yearc AS year, trend, timeperiodid AS month, 
flagname, region, geargrpcode,lon,lat, longitude,latitude,quadid AS quad, squaretypecode AS square, catchunit, dsettype, eff1, eff1type
FROM t2ce
WHERE region ='AT' AND yearc > 2009 AND timeperiodid < 13 AND eff1type IN ('D.FISH','FISH.HOUR') 
AND geargrpcode = 'BB' AND flagname IN ('Belize','Côte D''Ivoire','EU.España','Guatemala','Guinée Rep.','Panama','Venezuela','Panama','EU.France','EU.Portugal','U.S.A.', 'Curaçao','EU.Netherlands','Maroc','Namibia' ) 
AND catchunit != '--' ;")




```



```{r extract kg bigeye caught by selected flags from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE,eval=T}
bet <- sqlQuery(chan,"SELECT yearc AS year, trend,quadid as quad, squaretypecode AS square,
timeperiodid AS month, flagname, region, geargrpcode,lon,lat,longitude,latitude, 
catchunit, dsettype, eff1, eff1type, bet as measured_catch
FROM t2ce
WHERE region ='AT' AND timeperiodid < 13 AND yearc > 2009 AND 
geargrpcode IN ('LL','PS','BB') AND catchunit != '--' ;")
bet$species <- 'bet'


```



```{r extract kg skipjack caught by selected flags from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE,eval=T}
skj<- sqlQuery(chan,"SELECT yearc AS year, trend,quadid as quad, squaretypecode AS square,
timeperiodid AS month, flagname, region, geargrpcode,lon,lat,longitude,latitude, 
catchunit, dsettype, eff1, eff1type, skj as measured_catch
FROM t2ce
WHERE region ='AT' AND timeperiodid < 13 AND yearc > 2009 AND 
geargrpcode IN ('LL','PS','BB') AND catchunit != '--' ;")
skj$species <- 'skj'

```



```{r extract kg yellowfin caught by selected flags from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE,eval=T}
yft<- sqlQuery(chan,"SELECT yearc AS year, trend,quadid as quad, squaretypecode AS square,
timeperiodid AS month, flagname, region, geargrpcode,lon,lat,longitude,latitude, 
catchunit, dsettype, eff1, eff1type, yft as measured_catch
FROM t2ce
WHERE region ='AT' AND timeperiodid < 13 AND yearc > 2009 AND 
geargrpcode IN ('LL','PS','BB') AND catchunit != '--' ;")
yft$species <- 'yft'

```


#Task 2 data for long-line in the Atlantic


*Table 1. Effort type sampling by flag for longline in EFFDIS Task II database in 2014*
```{r make table for ll data by flag,include=TRUE,echo=FALSE}
pander(effort_type_by_flag_ll)
```



##Long-line effort (NO.HOOKS) distribution


```{r no of hooks shot by Japan in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(1,1))
three.d.effort.by.year(tdata=ll,which.year=2014,
                         which.flag='Japan',scaling.f=10000,effort.type='NO.HOOKS')  
```


**Figure 1.** *Number of hooks set by Japanese fleet (Task II only) in 2014*

```{r no of hooks shot by Chinese Taipei in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
three.d.effort.by.year(tdata=ll,which.year=2014,
                         which.flag="Chinese Taipei",scaling.f=10000,gridx=5,gridy=5,
                         effort.type='NO.HOOKS',which.gear='LL')  
```

**Figure 2.** *Number of hooks set by Chinese Taipei (Task II only) in 2014*



##Bigeye caught by long-line (2014).

```{r weight of bigeye caught in Atlantic in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,2),mar=c(2,1,2,1))
three.d.catch.by.year(tdata=bet,which.year=2014,which.gear='LL',gridx=5,gridy=5,which.species='bet',which.flag='Chinese Taipei',scaling.f=10,catchunit='kg')  
three.d.catch.by.year(tdata=bet,which.year=2014,
                        which.gear='LL',which.species='bet',which.flag='Japan',scaling.f=10,catchunit='nr')
three.d.catch.by.year(tdata=bet,which.year=2014,
                        which.gear='LL',which.species='bet',gridx=5,gridy=5,which.flag='EU.Portugal',scaling.f=10,catchunit='kg')  
three.d.catch.by.year(tdata=bet,which.year=2014,
                        which.gear='LL',which.species='bet',gridx=5,gridy=5,which.flag='Korea Rep.',scaling.f=10,catchunit='kg')
```

**Figure 3.** *Bigeye tuna caught by long-line by Chinese Taipei, Japan, EU. Portugal, and Korean Republic*


##Skipjack caught by long-line (2014).

```{r weight of skipjack caught by location by longline in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,1),mar=c(2,1,2,1))
three.d.catch.by.year(tdata=skj,which.year=2014,which.gear='LL',which.species='skj',which.flag='Chinese Taipei',scaling.f=1,catchunit='kg')  
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='LL',which.species='skj',which.flag='Japan',scaling.f=1,catchunit='nr')
```

**Figure 4.** *Skipjack tuna caught by long-line by Chinese Taipei, Japan*


##Yellowfin caught by long-line (2014).


```{r weight of yellowfin caught by location by longline in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(3,3),mar=c(2,2,2,2))
three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='LL',which.species='yft',which.flag='Chinese Taipei',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='LL',which.species='yft',which.flag='Japan',scaling.f=100,catchunit='nr')

three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='LL',which.species='yft',which.flag='U.S.A.',scaling.f=100,catchunit='nr')  
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='LL',which.species='yft',which.flag='Korea Rep.',scaling.f=100,catchunit='kg')

three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='LL',which.species='yft',which.flag='Canada',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='LL',which.species='yft',which.flag='China PR',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='LL',which.species='yft',which.flag='Mexico',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='LL',which.species='yft',which.flag='Vanuatu',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='LL',which.species='yft',which.flag='Venezuela',scaling.f=100,catchunit='kg')



```

**Figure 5.** *Yellowfin tuna caught by long-line by selected flags*


#Task 2 purse-seine data for the Atlantic


*Table 2. Effort type sampling by flag for purse-seiners in EFFDIS Task II database in 2014*
```{r make table for ps data by flag,include=TRUE,echo=FALSE}
pander(effort_type_by_flag_ps)
```




##Purse-seine effort (FISH.HOUR, D.FISH) distribution in the Atlantic (2014)


```{r no of fishing hours by Spain in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(1,1))
three.d.effort.by.year(tdata=ps,which.year=2014,
                         which.flag='EU.España',gridx=1,gridy=1,
                         scaling.f=2,effort.type='FISH.HOUR',which.gear='PS')  
```

**Figure 6.** *Hours fished reported by spanish purse-seiners (Task II only)  in 2014*

```{r no of fishing hours by Panama in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
three.d.effort.by.year(tdata=ps,which.year=2014,
                         which.flag='Panama',
                         gridx=1,gridy=1,scaling.f=1,effort.type='FISH.HOUR',which.gear='PS')  
```

**Figure 7.** *Hours fished reported by Panamanian purse-seiners (Task II only) 2014*


```{r no of fishing hours by Belize in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
three.d.effort.by.year(tdata=ps,which.year=2014,
                         which.flag='Belize',
                         gridx=1,gridy=1,scaling.f=1,effort.type='D.FISH',which.gear='PS')  
```

**Figure 8.** *Days fished reported by Belize purse-seiners (Task II only) 2014*


```{r no of fishing hours by Guatemala in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
three.d.effort.by.year(tdata=ps,which.year=2014,
                         which.flag='Guatemala',
                         gridx=1,gridy=1,scaling.f=1,effort.type='FISH.HOUR',which.gear='PS')  
```

**Figure 9.** *Hours fished reported by Belize purse-seiners (Task II only) 2014*



```{r no of fishing hours by Venezuela in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
three.d.effort.by.year(tdata=ps,which.year=2014,
                         which.flag="Venezuela",
                         gridx=1,gridy=1,scaling.f=1,effort.type='D.FISH',which.gear='PS')  
```

**Figure 10.** *Days fished reported by Venezuelan purse-seiners (Task II only) 2014*


```{r no of fishing hours by Curucao in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
three.d.effort.by.year(tdata=ps,which.year=2014,
                         which.flag="Curaçao",
                         gridx=1,gridy=1,scaling.f=1,effort.type='FISH.HOUR',which.gear='PS')  
```

**Figure 11.** *Hours fished reported by Curucao purse-seiners (Task II only) 2014*



```{r no of fishing hours by Guinee in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
three.d.effort.by.year(tdata=ps,which.year=2014,
                         which.flag="Guinée Rep.",
                         gridx=1,gridy=1,scaling.f=1,effort.type='FISH.HOUR',which.gear='PS')  
```

**Figure 12.** *Hours fished reported by Guinean purse-seiners (Task II only) 2014*


##Bigeye caught by purse-seine (2014).

```{r weight of bigeye caught by location by purse-seine in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,2),mar=c(3,3,3,3))
three.d.catch.by.year(tdata=bet,which.year=2014,which.gear='PS',gridx=1,gridy=1,which.species='bet',which.flag='EU.France',scaling.f=10,catchunit='kg')  
three.d.catch.by.year(tdata=bet,which.year=2014,
                        which.gear='PS',which.species='bet',gridx=1,gridy=1,which.flag='EU.España',scaling.f=10,catchunit='kg')
three.d.catch.by.year(tdata=bet,which.year=2014,
                        which.gear='PS',which.species='bet',gridx=1,gridy=1,which.flag='Venezuela',scaling.f=10,catchunit='kg')  
three.d.catch.by.year(tdata=bet,which.year=2014,
                        which.gear='PS',which.species='bet',gridx=1,gridy=1,which.flag='Panama',scaling.f=10,catchunit='kg')
```

**Figure 13.** *Bigeye tuna caught by purse seine by France, Spain, Venezuela, and Panama*

##Skipjack caught by purse-seine.

```{r weight of skipjack caught by location by purse-seine in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(3,3),mar=c(2,2,2,2))
three.d.catch.by.year(tdata=skj,which.year=2014,which.gear='PS',gridx=1,gridy=1,which.species='skj',which.flag='EU.France',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='PS',which.species='skj',gridx=1,gridy=1,which.flag='EU.España',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='PS',which.species='skj',gridx=1,gridy=1,which.flag='Venezuela',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='PS',which.species='skj',gridx=1,gridy=1,which.flag='Panama',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='PS',which.species='skj',gridx=1,gridy=1,which.flag='Belize',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='PS',which.species='skj',gridx=1,gridy=1,which.flag='Guatemala',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='PS',which.species='skj',gridx=1,gridy=1,which.flag="Côte D'Ivoire",scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='PS',which.species='skj',gridx=1,gridy=1,which.flag="Curaçao",scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='PS',which.species='skj',gridx=1,gridy=1,which.flag="Guinée Rep.",scaling.f=100,catchunit='kg')


```

**Figure 14.** *Skipjack caught by purse-seine by selected nations*

##Yellowfin caught by purse-seine.


```{r yellowfin caught by location by purse-seine in 2014 from task 2 part 1,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,2),mar=c(2,2,2,2))
three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='PS',gridx=1,gridy=1,which.species='yft',which.flag='EU.France',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='PS',which.species='yft',gridx=1,gridy=1,which.flag='EU.España',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='PS',which.species='yft',gridx=1,gridy=1,which.flag='Venezuela',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='PS',which.species='yft',gridx=1,gridy=1,which.flag='Panama',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='PS',which.species='yft',gridx=1,gridy=1,which.flag='Belize',scaling.f=100,catchunit='kg')
```

**Figure 15.** *Yellowfin caught by purse-seine by Spain, Venezuela, Panama, and Belize*



```{r yellowfin caught by location by purse-seine in 2014 from task 2 part 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,2),mar=c(2,2,2,2))
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='PS',which.species='yft',gridx=1,gridy=1,which.flag='Guatemala',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='PS',which.species='yft',gridx=1,gridy=1,which.flag="Côte D'Ivoire",scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='PS',which.species='yft',gridx=1,gridy=1,which.flag="Curaçao",scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='PS',which.species='yft',gridx=1,gridy=1,which.flag="Guinée Rep.",scaling.f=100,catchunit='kg')
```

**Figure 16.** *Yellowfin caught by purse-seine by Guatemala, Ivory Coast, Curacao, and Guinea*



#Task 2 bait-boat data for The Atlantic


*Table 3. Effort type sampling by flag for purse-seiners in EFFDIS Task II database in 2014*
```{r make table for bb data by flag,include=TRUE,echo=FALSE}
pander(effort_type_by_flag_bb)
```



##Bigeye caught by bait-boat (2014).


```{r weight of bigeye caught by location in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,3),mar=c(2,2,2,2))
three.d.catch.by.year(tdata=bet,which.year=2014,which.gear='BB',gridx=1,gridy=1,which.species='bet',which.flag='EU.France',scaling.f=10,catchunit='kg')  
three.d.catch.by.year(tdata=bet,which.year=2014,
                        which.gear='BB',which.species='bet',gridx=1,gridy=1,which.flag='EU.España',scaling.f=10,catchunit='kg')
three.d.catch.by.year(tdata=bet,which.year=2014,
                        which.gear='BB',which.species='bet',gridx=1,gridy=1,which.flag='Venezuela',scaling.f=10,catchunit='kg')  
three.d.catch.by.year(tdata=bet,which.year=2014,which.gear='BB',which.species='bet',gridx=1,gridy=1,which.flag='South Africa',scaling.f=10,catchunit='kg')

three.d.catch.by.year(tdata=bet,which.year=2014,which.gear='BB',which.species='bet',gridx=1,gridy=1,which.flag='Senegal',scaling.f=10,catchunit='kg')

```

**Figure 17.** *Bigeye caught by selected nations by bait-boat in 2014*


##Skipjack caught by bait-boat in 2014.



```{r skipjack caught by location in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,3),mar=c(2,2,2,2))
three.d.catch.by.year(tdata=skj,which.year=2014,which.gear='BB',gridx=1,gridy=1,which.species='skj',which.flag='EU.France',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='BB',which.species='skj',gridx=1,gridy=1,which.flag='EU.España',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=skj,which.year=2014,
                        which.gear='BB',which.species='skj',gridx=1,gridy=1,which.flag='Senegal',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=skj,which.year=2014,which.gear='BB',which.species='skj',gridx=1,gridy=1,which.flag='South Africa',scaling.f=10,catchunit='kg')

three.d.catch.by.year(tdata=skj,which.year=2014,which.gear='BB',which.species='skj',gridx=1,gridy=1,which.flag='Venezuela',scaling.f=10,catchunit='kg')

three.d.catch.by.year(tdata=skj,which.year=2014,which.gear='BB',which.species='skj',gridx=1,gridy=1,which.flag='Brazil',scaling.f=10,catchunit='kg')

```

**Figure 18.** *Skipjack caught by bait-boat by 6 nations in 2014*


##Yellowfin caught by bait-boat in 2014.

```{r yellowfin caught by baitboat by location in 2014 from task 2,include=TRUE,echo=F,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,3),mar=c(2,2,2,2))
three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='BB',gridx=1,gridy=1,which.species='yft',which.flag='EU.France',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='BB',which.species='yft',gridx=1,gridy=1,which.flag='EU.España',scaling.f=100,catchunit='kg')
three.d.catch.by.year(tdata=yft,which.year=2014,
                        which.gear='BB',which.species='yft',gridx=1,gridy=1,which.flag='Senegal',scaling.f=100,catchunit='kg')  
three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='BB',which.species='yft',gridx=1,gridy=1,which.flag='South Africa',scaling.f=100,catchunit='kg')

three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='BB',which.species='yft',gridx=1,gridy=1,which.flag='Venezuela',scaling.f=100,catchunit='kg')

three.d.catch.by.year(tdata=yft,which.year=2014,which.gear='BB',which.species='yft',gridx=1,gridy=1,which.flag='Brazil',scaling.f=100,catchunit='kg')


```

**Figure 19.** *Yellowfin tuna caught by bait-boat by 6 nations*






























