---
title: "Distribution of long-line, purse-seine and baitboat fishing activity in the Atlantic Ocean"
author: "Doug Beare"
date: "3 December 2015"
output: html_document
header-includes: \usepackage{tabularx}
subtitle: null
Grmd::docx_document:
  fig_caption: yes
  force_captions: yes
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

#Objective of Atlantic Tuna Tagging Programme (AOTTP) 

The overall objective of the Atlantic Tuna Tagging Programme (AOTTP) is to contribute to food security and economic growth of the Atlantic developing coastal states by ensuring sustainable management of tropical tuna resources in the Atlantic Ocean
The specific objective of this programme is to provide evidence based scientific advices to developing costal states, and other Contracting Parties, to support the adoption of effective Conservation and Management Measures (CMMs) in the framework of the International Commission for the Conservations of Atlantic Tunas (ICCAT). This will be achieved through improving the estimation, derived from tag-recapture data, of key parameters for stock assessment analyses, i.e. growth, natural mortality, movements and stock structure, etc.

#Task 2 catch and effort data exploration for AOTTP

##Install required r libraries

Install, e.g. install.packages("spatial") and 'attach' the R-libraries required. 

```{r attach libraries, include =TRUE, echo=T, message=FALSE,error=FALSE,warning=FALSE}
library(spatial)
library(sp)
library(doBy)
library(rgdal)
library(RODBC)
library(RColorBrewer)
library(ggplot2)
library(vmstools)
library(mgcv)
library(maps)
library(mapdata)
library(reshape2)
library(rgeos)
library(lattice)
library(pander)
library(kfigr)
library(doMC)
```

##Install *effdisR*
Install and load the R-scripts and data contained in the (*effdisR*) R-package. The tarball can be obtained here, https://github.com/bearedo/effdis/blob/master/effdisR/effdisR\_1.0.tar.gz.  It can be installed in R as follows:

```{r install effdis R library,eval=FALSE,include=TRUE,echo=TRUE}
install.packages('/path/to/package', repos = NULL, type="source")
```

```{r load effdis R library,eval=TRUE,include=TRUE,echo=TRUE}
library(effdisR)
```

##Connect to the Task II database


```{r connect to db,include=TRUE,echo=TRUE}
chan <- odbcConnect("effdis-tuna-cc1", case="postgresql", believeNRows=FALSE)
```

'chan' is then stored in R and allows you to connect to the remote database, download data etc.

##Exploratory data analysis

```{r get ll data by flag,include=TRUE,echo=TRUE}
effort_type_by_flag_ll <- sqlQuery(chan, "SELECT flagname AS Flag, 
eff1type AS Effort_type, count(eff1type) as No_records
                 FROM t2ce_new
                 WHERE region ='AT' AND geargrpcode = 'LL' AND yearc = 2014
                 GROUP BY flagname, eff1type
                 ORDER BY flagname, eff1type, No_records;")
```


*Table 1. Effort type sampling by flag for longline in EFFDIS Task II database in 2014*
```{r make table for ll data by flag,include=TRUE,echo=FALSE}
pander(effort_type_by_flag_ll)
```


```{r get ps data by flag,include=TRUE,echo=TRUE}
effort_type_by_flag_ps <- sqlQuery(chan, "SELECT flagname AS Flag, 
eff1type AS Effort_type, count(eff1type) as No_records
                 FROM t2ce_new
                 WHERE region ='AT' AND geargrpcode = 'PS' AND yearc = 2014
                 GROUP BY flagname, eff1type
                 ORDER BY flagname, eff1type, No_records;")
```

*Table 2. Effort type sampling by flag for purse-seiners in EFFDIS Task II database in 2014*
```{r make table for ps data by flag,include=TRUE,echo=FALSE}
pander(effort_type_by_flag_ps)
```


```{r get bb data by flag,include=TRUE,echo=TRUE}
effort_type_by_flag_bb <- sqlQuery(chan, "SELECT flagname AS Flag, 
eff1type AS Effort_type, count(eff1type) as No_records
                 FROM t2ce_new
                 WHERE region ='AT' AND geargrpcode = 'BB' AND yearc = 2014
                 GROUP BY flagname, eff1type
                 ORDER BY flagname, eff1type, No_records;")
```

*Table 3. Effort type sampling by flag for purse-seiners in EFFDIS Task II database in 2014*
```{r make table for ps data by flag,include=TRUE,echo=FALSE}
pander(effort_type_by_flag_bb)


### Data coverage

It is important to understand how the distribution of samples in the Task II database varies with respect to location and time (long-term and seasonal). The function *yr.month.coverage.task2* available in *effdisR* counts the number of samples by year and month for any strata (gear, flag etc) and displays the results as a 3D plot. This type of plot reveals non-random sampling in time.  It is possible, for example, that sampling might have concentrated on the first part of the year for a decade, and then switched to the latter part of the year. 'Trends' estimated from data collected in such a manner will clearly be spurious. Examples of the output of *yr.month.coverage.task2* are plotted in Figure 2 for longline between 1960 and 2010 for Japan, Chinese Taipei, Brasil, and U.S.A. Clearly the extent of the data available varies substantially between flags. There are no obvious seasonal biases in the data but the amount of reporting has changed with long-term time. Japan has been particularly consistent (Fig. 2, top right), while the U.S.A. has been inconsistent.  

```{r get ll data for ten most important flags,include=TRUE,echo=TRUE}
ll <- sqlQuery(chan,"SELECT yearc AS year, trend, timeperiodid AS month, 
flagname, region, geargrpcode,longitude,latitude, catchunit, dsettype, eff1, eff1type
FROM t2ce_new
WHERE region ='AT' AND yearc > 2009 AND timeperiodid < 13 AND eff1type='NO.HOOKS' AND geargrpcode = 'LL' 
AND flagname IN ('Japan','Chinese Taipei','Brazil','Angola','EU.España','EU.Portugal','St. Vincent and Grenadines','Vanuatu','U.S.A.','Namibia') 
AND catchunit != '--' ;")
```


###Sampling in space by year

The distribution of data/samples in space is similarly important. The function *spatial.coverage.by.year.task2* plots the distribution of Task II data by location for any combination of flag, gear etc. Output is illustrated for longliners for two arbitarily selected flags and time-periods in Figures 3 and 4.

```{r spatial sampling Brazil,include=TRUE,echo=TRUE}
par(mfrow=c(2,2),mar=c(0,0,1,0))
spatial.coverage.by.year.task2(tdata=ll,
        start.year=2011,end.year=2014,which.gear='LL',which.flag='Brazil')
```

**Figure 3.** *Spatial distribution data: Brazilian long-liners between 2011 and 2014*

The extent of Brazilian longlining activity has been fairly consistent between 2011 and 2014 (Fig. 4). In contrast fleets from both Japan and Chinese Taipei cover substantial areas of the Atlantic Ocean (Figs. 5 and 6).

```{r spatial sampling Chinese Taipei, include=TRUE,echo=TRUE}
par(mfrow=c(2,2),mar=c(0,0,1,0))
spatial.coverage.by.year.task2(tdata=ll,
        start.year=2011,end.year=2014,which.gear='LL',which.flag='Chinese Taipei')
```

**Figure 4.** *Spatial sampling by Chinese Taipei long-liners between 2011 and 2014*


```{r spatial sampling EU. Spain, include=TRUE,echo=TRUE}
par(mfrow=c(2,2),mar=c(0,0,1,0))
spatial.coverage.by.year.task2(tdata=ll,
        start.year=2011,end.year=2014,which.gear='LL',which.flag='EU.España')
```

**Figure 5.** *Spatial sampling by EU. Spain long-liners between 2011 and 2014*

```{r spatial sampling St Vincent and Grenadines, include=TRUE,echo=TRUE}
par(mfrow=c(2,2),mar=c(0,0,1,0))
spatial.coverage.by.year.task2(tdata=ll,
        start.year=2011,end.year=2014,which.gear='LL',which.flag="St. Vincent and Grenadines")
```

**Figure 6.** *Spatial sampling by St Vincent and Grenadines long-liners between 2011 and 2014*


```{r spatial sampling St Vincent and Grenadines, include=TRUE,echo=TRUE}
par(mfrow=c(2,2),mar=c(0,0,1,0))
spatial.coverage.by.year.task2(tdata=ll,
        start.year=2011,end.year=2014,which.gear='LL',which.flag="U.S.A.")
```

**Figure 7.** *Spatial sampling by USA long-liners between 2011 and 2014*





```{r get purse-seine data,include=TRUE,echo=TRUE,fig.height=5}
ps <- sqlQuery(chan,"SELECT yearc AS year, trend, timeperiodid AS month, 
flagname, region, geargrpcode,longitude,latitude, catchunit, dsettype, eff1, eff1type
FROM t2ce
WHERE region ='AT' AND yearc > 2004 AND timeperiodid < 13 AND eff1type IN ('D.FISH','FISH.HOUR') 
AND geargrpcode = 'PS' AND flagname IN ('Belize','Côte D''Ivoire','EU.España','Guatemala','Guinée Rep.','Panama') 
AND catchunit != '--' ;")
```

```{r plot distribution of spanish purse-seine data,include=TRUE,echo=TRUE,fig.height=5}
par(mfrow=c(2,2),mar=c(1,1,2,1),oma=c(3,3,3,3))
spatial.coverage.by.year.task2(tdata=ps,start.year=2005,end.year=2008,
                                 which.gear='PS',which.flag='EU.España')
```

**Figure 5.** *Spatial sampling by Spanish purse-seiners between 1995 and 2010*

Most of the purse-seining data have been supplied at a higher spatial resolution (1x1) than is available for the long-liners. This is illustrated in Figure 5 which shows the distribution of Spanish purse-seining activity off the West Coast of Africa between 2005 and 2010.

### Effort by year and location (raw data, no modeling)

Given that we can now determine the timing and location of fishing activities (Figs. 2-6) we now need to know its intensity, ie. how many hooks were set or days fished at a particular location ? Task II effort data of any type can be plotted spatially using the R function *three.d.effort.by.year* and example output is shown in Figures 6 to 9. 

In 2006, for example, longlining effort by Japan focused on the North and Eastern Atlantic (Fig. 6). Chinese Taipei flagged long-liners, on the other hand also worked in the North Atlantic but their effort appears greater in the South Atlantic (Fig. 7). Purse-seining data are submitted to ICCAT at a higher spatial resolution (1x1 grid) and examples for Spain and Panama in 2009 are plotted in Figures 8 and 9. Both fleets focus their activity along the West Coast of Africa but the Panamanian fleet tends not to venture as far inshore as the Spanish fleet (Figs 8 & 9).

```{r no of hooks shot by Japan in 2006 from task 2,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(1,1))
three.d.effort.by.year(tdata=ll,what.year=2006,
                         what.flag='Japan',scaling.f=10000,effort.type='NO.HOOKS')  
```

**Figure 6.** *Total number of hooks set by Japanese fleet (Task II only) in 2006*

```{r no of hooks shot by Chinese Taipei in 2006 from task 2,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
three.d.effort.by.year(tdata=ll,what.year=2006,
                         what.flag="Chinese Taipei",scaling.f=10000,gridx=5,gridy=5,
                         effort.type='NO.HOOKS',what.gear='LL')  
```

**Figure 7.** *Total number of hooks set by Chinese Taipei (Task II only) in 2006*

```{r no of fishing hours by Spain in 2009 from task 2,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
three.d.effort.by.year(tdata=ps,what.year=2009,
                         what.flag='EU.España',gridx=1,gridy=1,
                         scaling.f=2,effort.type='FISH.HOUR',what.gear='PS')  
```

**Figure 8.** *Total fishing hours reported by Spanish purse-seiners (Task II only)  in 2009*

```{r no of fishing hours by Other in 2009 from task 2,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
three.d.effort.by.year(tdata=ps,what.year=2009,
                         what.flag='Panama',
                         gridx=1,gridy=1,scaling.f=1,effort.type='FISH.HOUR',what.gear='PS')  
```

**Figure 9.** *Total fishing hours reported by Panamanian purse-seiners (Task II only) 2009*


### Catch weights by year and location - Japanese longliners, and Spanish Purse-seiners

With *circa* 27 flags, multiple species, 12 months, 61 years, and different effort submissions (e.g. days at sea, number of hooks set) there is a large number of possible combinations for examining the catches. Examples are illustrated in Figure 10 using the R function, *three.d.catch.by.year* linked to the PostgreSQL database via an SQL script (see below). In this example we extracted data for albacore tuna caught by long-line for Japan, and by purse-seine for Spain. [Note that observations were ignored where the catchunit is unknown ('--'). Numbers or kilograms caught can be selected depending on availability].


```{r extract kg tunas caught by selected flags in 2006 from task 2,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
alb <- sqlQuery(chan,"SELECT yearc AS year, trend, 
timeperiodid AS month, flagname, region, geargrpcode,longitude,latitude, 
catchunit, dsettype, eff1, eff1type, alb as measured_catch
FROM t2ce
WHERE region ='AT' AND timeperiodid < 13 AND eff1type IN ('NO.HOOKS','FISH.HOUR') AND 
geargrpcode IN ('LL','PS') AND flagname IN ('Japan','EU.España') AND catchunit != '--' ;")
alb$species <- 'alb'
```


```{r weight of albacore caught by location in 2006 from task 2,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(2,2),mar=c(2,1,2,1))
three.d.catch.by.year(tdata=alb,what.year=2006,what.gear='LL',what.species='alb',what.flag='Japan',scaling.f=10,catchunit='nr')  
three.d.catch.by.year(tdata=alb,what.year=2009,
                        what.gear='LL',what.species='alb',what.flag='Japan',scaling.f=10,catchunit='nr')
three.d.catch.by.year(tdata=alb,what.year=1992,
                        what.gear='PS',what.species='alb',gridx=1,gridy=1,what.flag='EU.España',scaling.f=25,catchunit='kg')  
three.d.catch.by.year(tdata=alb,what.year=1995,
                        what.gear='PS',what.species='alb',gridx=1,gridy=1,what.flag='EU.España',scaling.f=25,catchunit='kg')
```

**Figure 10.** *Weight of albacore tuna caught by longline by Japan (2006 & 2009) and purse-seine by Spain (1992 & 1995)*

# EFFDIS effort estimation with *effdisR*: worked example (Japanese longline)

##Step 1 
Get Task II data from the database for each dsettype using *get.effdis.t2.data* and then combine them using rbind. 

```{r get japanese task 2 longline data,eval=FALSE,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
ll_n  <- get.effdis.t2.data(which.dsn='effdis-tuna-cc1',
                              which.gear='LL',which.flag='All',which.dsettype = 'n-')
ll_nw <- get.effdis.t2.data(which.dsn='effdis-tuna-cc1',
                              which.gear='LL',which.flag='All',which.dsettype = 'nw')
ll_w  <- get.effdis.t2.data(which.dsn='effdis-tuna-cc1',
                              which.gear='LL',which.flag='All',which.dsettype = '-w')
long_line <- rbind(ll_n,ll_nw,ll_w)
```

##Step 2
Find and extract those EFFDIS data that are in the Atlantic using *find.ocean*. This function uses a shapefile which is included in *effdisR*, and can be examined by typing *data("seas")* and then *plot(seas)*.  The definitions of 'Atlantic' used by ICCAT are rather crude. Thus function tries to mimic ICCAT definitions and 'Atlantic', for example, includes the Gulf of St Lawrence, the Caribbean Sea, The Celtic Sea, The Gulf of Guinea, The Gulf of Mexico and The Irish Sea. Obviously this is flexible and can easily be changed.  Data for land, the Mediterranean and Pacific can be extracted using the strings, 'land', 'med', and 'pac' respectively in the place of 'atl' below. 

```{r identify data in Atlantic,eval=FALSE,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
data("seas")
long_line<-find.ocean(long_line)
long_line <- long_line[long_line$which.ocean == 'atl',]
```

##Step 3
Make sure the data are 'clean' using *prepare.effdis.data* as follows. Sometimes downloading from a database creates 'factor' objects in R which can cause problems. This function converts them to character strings.

```{r get Japanese task 2 longline data,eval=FALSE,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
long_line<-prepare.effdis.data(input=long_line)
```

##Step 4
Convert data from 'short format' to 'long format' using *convert2long.format.t2*. We do this because the long format is required for subsequent regression modeling.

```{r Convert data to long format,eval=FALSE,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
long_line_lf <- convert2long.format.t2(input =long_line)
```

##Step 5
Estimate catch weights from numbers where none are supplied using the functions *model.nos.kgs* (fits the model) and *kgs.from.nos* which imputes weights for countries that supply only numbers caught. Note that some countries report catches by total weight, some by total numbers and some by both. For the purposes of the effort estimations we are attempting to make here, it is essential that the catch data are available in the same unit of measurement.  The data for countries that have reported both weights and numbers were, therefore, extracted and examined together. The relationships betweem them are highly linear (see Interim Report) and we decided to model the weight caught as a (linear) function of number caught plus other useful, predictive covariates (e.g. flag, species, and trend). A stepwise model selection procedure selects the 'best' model (bm) which is then used to impute catch weights in kgs for Task II in cases where total numbers only were supplied, e.g. U.S.A. The model below (bm) fits the data well and explains most of the variance ($R^2$ = 83%).  This part of the procedure/model is included in the error variance estimation. 

```{r model weights as a function of numbers,eval=FALSE,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
bm <- model.nos.kgs(input=long_line_lf,which.gear='LL')
long_line_lf <- kgs.from.nos(long_line_lf) 
```

```{r load best.model ,eval=TRUE,include=FALSE,echo=FALSE,message = FALSE, error=FALSE, warnings=FALSE}
load('/home/doug/effdis/data/best.model.ll.RData')
```

*Table 3. Linear model summarising the relationship between weights and numbers for countries that sent both to ICCAT for the Task II database*
```{r display bestmodel,eval=TRUE,include=TRUE,echo=TRUE}
panderOptions("digits",1)
pander(bm)
```

##Step 6
Fit regression models to the Task II catch and effort data using *fit2stageGAMtoCatch* and *fitGAMtoEffort*. This step is illustrated here for speed and convenience with bluefin tuna only ('bft') and only between 1990 and 2000. To get a global estimate of effort, however, you would, of course, need to do the same for all nine species but it would be straightforward to add any others if necessary.  The functions fit and test a suite of generalised additive models (GAMs) fitted to the Task II catch and effort data.  GAMs were selected because they are highly flexible, impose no particular functional form on the data, and they can deal with skew distributions and high prevalences of zeros.  The models take the relevant variables (eg. number of hooks set, weight of fish caught) and model them as smooth functions of various combinations of covariates of location (latitude, longitude, bottom depth) and time (year and month). 

The first stage models the probability of recording a catch using a GAM from the Quasibinomial family (Model 1), where $P$ is the probability of catching a fish.

1. $P_{xytm} = s(x,y)+s(t)+s(m)+\epsilon$

We then model the positive component of the catch, $C$, with a GAM from the Gamma family (Model 2).

2. $C_{xytm} = s(x,y)+s(t)+s(m)+\epsilon$

And finally the fishing effort (number of hooks) is modeled using a GAM from the QuasiPoisson family (Model 3).

3. $E_{xytm} = s(x,y)+s(t)+s(m)+\epsilon$

In all 3 models, $x$,$y$,$t$, and $m$ are longitude, latitude, trend, and month respectively and $s$ are spline smooth functions fitted by generalised-cross-validation using the MGCV R-library, see https://cran.r-project.org/web/packages/mgcv/mgcv.pdf. $\epsilon$ is different, as appropriate, for each model. 

```{r model catches as a function of space and time,eval=FALSE, include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
j_bft_ll <- fit2stageGAMtoCatch(input=long_line_lf,
                                  which.flag='Japan',which.species='bft',start.year=1970,end.year=2010)
j_emod_ll <- fitGAMtoEffort(input=long_line_lf,
                              which.flag='Japan',which.effort='NO.HOOKS',start.year=1970,end.year=2010)
```

##Step 7
Use the models to predict values over a grid of 'new data'.  Currently this takes the range of locations **ever** recorded in the data, constructs a grid for each time-step (1950 to present by month) and makes the predictions with the model. The function *predict.effdis.t2.data* also identifies and flags up those points in space and time where data were actually collected.

```{r predict catches over grid,eval=FALSE,include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
j_bft.ll.pred <- predict.effdis.t2.data(cmod=j_bft_ll,
                                          effmod=j_emod_ll,grid.res=5,
                                          start.year=1970,end.year=2010,which.flag='Japan')
```

The fitted model parameters are used to 'predict' values of catch, effort and catch-per-unit-effort on a grid of all combinations of the selected covariates, together with error or variance if required using the function *predict.effdis.t2.data*.  Note that total Task II catch is calculated by multiplying the predictions from Models 1 and 2, ie. 'given that fish were caught, how many/much?', in a 'two-stage process'.  The function *plot.mods* can be used to plot selected subsets of the model output. A map of the 'probability of the Japanese long-line fleet catching a blue fin tuna in January 1995 can, for example, be created using the following code:

```{r load predictions for Japanese bluefin long-line,eval=TRUE,include=FALSE,echo=FALSE,message = FALSE, error=FALSE, warnings=FALSE}
load('/home/doug/effdis/data/j_bft_ll.RData')
load('/home/doug/effdis/data/j_bft.ll.pred.RData')
```

```{r plot P over grid,eval=TRUE, include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
par(mfrow=c(1,2))
plot.mods(input=j_bft.ll.pred,cmod=j_bft_ll,
            which.year = 1995,which.month=1,which.value = 'prob',grid.res=5,which.gear='LL')
plot.mods(input=j_bft.ll.pred,cmod=j_bft_ll,
            which.year = 1995,which.month=1,which.value = 'prob',grid.res=5,which.gear='LL',
            plot.samples.only = FALSE)
```

**Figure 11.** *Probability of catching a bluefin tuna (Japanese longline fleet)*

Figure 11 shows predictions for bluefin tuna from the binary (Bernouilli) model for January 1995. The probability of catching one is highest in the Caribbean, Gulf of Mexico and North Atlantic.  The left-hand plot shows the model output for grid cells where a real observation exists, while the right-hand plot is the interpolation based on the area of the entire dataset.

## Step 8
Obtain (and aggregate for later use) Task I long-line data from the database using *get.effdis.t1.data*. Note that Task I are annual catch totals which are thought to be comprehensive.

```{r get task 1 data,eval=FALSE, include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
long_line.t1 <- get.effdis.t1.data(which.dsn='effdis-tuna-cc1',
                                     which.gear = 'LL',which.region='AT',which.flag='Japan')
long_line.sum.t1 <- aggregate(list(qty_t=long_line.t1$qty_t),
                              list(year=long_line.t1$yearc),sum,na.rm=T)
```

##Step 9
Collect model data together into a single database. In our method you bind up the model estimates for all 9 species as follows. Remember that you need to run Steps 6 and 7 for each species (code not shown here).

```{r bind task 2 estimates ,eval=FALSE, include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
model.data <- rbind(j_alb.ll.pred,j_bft.ll.pred,j_bet.ll.pred,
                    j_bum.ll.pred,j_skj.ll.pred,j_yft.ll.pred,
                    j_swo.ll.pred,j_sai.ll.pred,j_whm.ll)
```

##Step 10
Block out places and times for which data were never collected. This is optional and if all the data are used (see Figure 11 above) the results are very similar.

```{r block out unsampled grids ,eval=FALSE, include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
model.data$catch[big$observation == FALSE] <- NA
model.data$prob[big$observation == FALSE] <- NA
model.data$measured_catch[big$observation == FALSE] <- NA
model.data$eff[big$observation == FALSE]  <- NA
```

##Step 11
Convert Task II catches from kgs to tonnes to match with the Task I data.

```{r convert task 2 to tonnes ,eval=FALSE, include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
model.data$catch <- model.data$catch/1000
```

##Step 12
Sum Task II data (catch and effort) for all 9 species.

```{r sum modeled task 2 over year ,eval=FALSE, include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
model.data.totals <- aggregate(list(catch=model.data$catch,catch=model.data$catch,eff=model.data$eff), 
                  by=list(year=model.data$year),sum,na.rm=T)
```

##Step 13
Merge the Task I and modeled Task II totals, calculate a global, modeled Task II CPUE and raise by Task I CPUE to give raised effort.

```{r merge task 1 and task 2 totals,eval=FALSE, include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
t1.t2.merged <- merge(model.data.totals,long_line.sum.t1)
t1.t2.merged$cpue <- t1.t2.merged$catch/t1.t2.merged$eff
t1.t2.merged$raised.effort <- t1.t2.merged$qty_t/t1.t2.merged$cpue
```

##Step 14
Plot the raised effort as a function of year.

```{r plot estimate for Japan 1970-2010,eval=TRUE, include=TRUE,echo=TRUE,message = FALSE, error=FALSE, warnings=FALSE}
data("japan-effdis-estimates")
par(mfrow=c(1,1),mar=c(5,5,4,4))
plot(effdis$year,effdis$raised.effort/1000000,xlab='',ylab='hooks',type='l',lwd=3,xlim=c(1970,2010))
abline(v=seq(1960,2010,by=5),lty=2,col='blue')
title('Task I Catch / Task II CPUE for Japanese long-liners')
```

**Figure 12.** *Estimate of total effort (no of hooks) calculated for Japanese long-liners according to $C_{taskI}/U_{taskII}$ where C=catch and U = catch-per-unit effort.*


#Conclusions

The use of the *effdisR* library for raising Task II EFFDIS effort estimates by Task I totals has been demonstrated here for Japanese longliners. The code above also works for other gears (purse-seiners, baitboats) but obviously different input parameters need to be inserted into the functions. The estimate shown here for the Japanese long-line fleet is very similar to that calculated by de Bruyn et al (2014). To get global estimates the code above should be re-run for each fleet or flag and the estimates summed. 































