---
title: ' EFFDIS Effort Estimation'
author: "Doug Beare, Laurie Kell & Carlos Palma"
date: "03/06/2015"
output: slidy_presentation
---

## SHORT-TERM CONTRACT: A MODELING APPROACH TO ESTIMATE OVERALL ATLANTIC FISHING EFFORT BY TIME-AREA STRATA (EFFDIS)

### Objectives

* Develop a robust statistical modeling approach to estimate overall Atlantic fishing effort stratified by flag/fleet, gear, area (5ºx5º degree square grid), year and month (starting in 1950).

* Update the current EFFDIS estimations for longline gear (1950 to 2014) using the new approach, and then develop estimation procedures for baitboat and purse-seine with the appropriate effort units.  

## The SGECO Working group made a series of recommendations for improving EFFDIS, ie.to:

+ Consider seasonal and spatial patterns and their interactions 
+ Understand how information from species composition can best be used in this context
+ Combine bait boat and purse-seine estimates with long-line
+ Estimate uncertainty/variance 
+ Exploit other relevant information where available, e.g. VMS data

## Overall workplan

+ Obtain all the relevant data and review the current methods for EFFDIS estimation 
+ Write documented R code to mimic the current procedure
+ Develop  ‘strawman’ methodologies for estimating fishing effort for a single fleet/flag country
+ Once a method is approved for one fleet the Contractor will then adapt it to other fleets to produce global estimates  
+ Use an online SQL relational database linked to R-scripts

## Detailed workplan

+ Explore data for non-random, non-representative sampling (data catalogue)
+ Investigate the relationship between Task 1 and Task 2 data (Sum of products) for fleet combinations
+ Develop most appropriate multi-variate models (e.g. GLMs, GAMs) for interpoplation
+ Estimate variance/bias with jack-knife
+ Test predictions using cross-validation methodology

## Exploring data for non-random, non-representative sampling (data catalogue)

### Sampling by year and month - Chinese Taipei

```{r code,include=FALSE}
library(rio)
library(maps)
library(mapdata)
library(lattice)
source("/home/doug/effdis/R/yr.month.coverage.task2.r")
source("/home/doug/effdis/R/spatial.coverage.by.year.task2.r")
t2ceLL <- read.table("/home/doug/effdis/data/t2ceLL.csv",sep=",")
t1det9sp <- import("/home/doug/effdis/data/t1det_9sp.xlsx")
```
```{r temporal confounding Chinese Taipei,include=TRUE,echo=FALSE}
yr.month.coverage.task2.r(tdata=t2ceLL,which.flag='Chinese Taipei')
```

## Sampling by year and month - Japan

```{r temporal confounding Japan,include=TRUE,echo=FALSE}
yr.month.coverage.task2.r(tdata=t2ceLL,which.flag='Japan')
```

## Sampling in space by year - Chinese Taipei 

```{r spatial confounding China,include=TRUE,echo=FALSE}
library(maps)
spatial.coverage.by.year.task2.r(tdata=t2ceLL,which.flag='Chinese Taipei')
```

## Sampling in space by year - Portugal

```{r spatial confounding by month Portugal,include=TRUE,echo=FALSE}
source("/home/doug/effdis/R/spatial.coverage.by.month.task2.r")
spatial.coverage.by.month.task2.r()
```

## No of hooks by time - Chinese Taipei and Japan

```{r no of hooks reported by Taipei and Japan in task 2,include=TRUE,echo=FALSE,message = FALSE, error=FALSE, warnings=FALSE}
source("/home/doug/effdis/R/effort.by.year.task2.r")
par(mfrow=c(2,1))
effort.by.year.task2.r(which.flag='Chinese Taipei')
effort.by.year.task2.r(which.flag='Japan')
```


## Relationship between Task 1 and Task 2 data (Sum of products) for all fleet combinations


```{r relationship between task 1 and task 2 data, include=TRUE,echo=FALSE,message = FALSE, error=FALSE, warnings=FALSE}
task1.2 <- read.table("/home/doug/effdis/data/task1.2.csv",header=T,sep=',')
xyplot(log(total_catch_kgs) ~ log(measured_catch)|flagname,groups=catchunit,data=task1.2[task1.2$region == 'AT' ,]
       ,xlab='task2',ylab='task1',auto.key = T)
```

## Developing multivariate models. Swordfish modeled (GAM) as function of space and time - Chinese Taipei for June 2009


```{r regression modeling, include=TRUE,echo=FALSE,message = FALSE, error=FALSE, warnings=FALSE}
grd <- read.table(file='/home/doug/effdis/R/grd')
lonnie <- seq(-87.5,22.5,by=5)
lattie <- seq(-57.5,52.5,by=5)
lo<- length(lonnie)
la<- length(lattie)

par(mfrow=c(2,2),mar=c(1,1,3,1))

image(lonnie,lattie,matrix(grd$hks,lo,la),col=topo.colors(100),xlab='',ylab='')
contour(lonnie,lattie,matrix(grd$hks,lo,la),col=topo.colors(100),add=T)
map('worldHires',add=T,fill=T);title('Number of hooks')


image(lonnie,lattie,matrix(grd$prob,lo,la),col=topo.colors(100),xlab='',ylab='')
contour(lonnie,lattie,matrix(grd$prob,lo,la),col=topo.colors(100),add=T)
map('worldHires',add=T,fill=T);title('Probability of catch (P)')


image(lonnie,lattie,matrix(grd$pcatch,lo,la),col=topo.colors(100),xlab='',ylab='')
contour(lonnie,lattie,matrix(grd$catch,lo,la),col=topo.colors(100),add=T)
map('worldHires',add=T,fill=T);title('Catch without zeros (C1)')


image(lonnie,lattie,matrix(grd$catch,lo,la),col=topo.colors(100),xlab='',ylab='')
contour(lonnie,lattie,matrix(grd$catch,lo,la),col=topo.colors(100),add=T)
map('worldHires',add=T,fill=T);title('Catch - P x C1')

```

## Conclusions
+ The project only recently started and the data are complex and the Contractor is working to understand the nuances
+ There is non-random, non-representative sampling
+ There is a strong positive correlation between Task 1 and Task 2 data for most countries which is useful information
+ Provision of numbers for Task 2 (instead of weights) by e.g. USA creates problems
+ Regression modeling is a promising approach
+ A series of models can be fitted to the data and used for interpolation and variance estimation 











