---
title: "ESTIMATING ATLANTIC FISHING EFFORT BY TIME-AREA STRATA"
author: "Doug Beare with assistance from the ICCAT Secretariat"
date: "24 July 2015"
output: beamer_presentation
---

# SHORT-TERM CONTRACT: A MODELING APPROACH TO ESTIMATE OVERALL ATLANTIC FISHING EFFORT BY TIME-AREA STRATA (EFFDIS)

## Objectives

* Develop a robust statistical modeling approach to estimate overall Atlantic fishing effort stratified by flag/fleet, gear, area (5ºx5º degree square grid), year and month (starting in 1950).

* Update the current EFFDIS estimations for longline gear (1950 to 2014) using the new approach, and then develop estimation procedures for baitboat and purse-seine with the appropriate effort units.  

## The SGECO Working group made a series of recommendations for improving EFFDIS, ie.to:

+ Consider seasonal and spatial patterns and their interactions 
+ Understand how information from species composition can best be used in this context
+ Combine bait boat and purse-seine estimates with long-line
+ Estimate uncertainty/variance 
+ Exploit other relevant information where available, e.g. VMS data, depth, sea temperature, primary production

## Overall workplan

+ Obtain all the relevant data and review the current methods for EFFDIS estimation 
+ Write documented R code to mimic the current procedure
+ Develop  ‘strawman’ methodologies for estimating fishing effort for a single fleet/flag country
+ Once a method is approved for one fleet the Contractor will then adapt it to other fleets to produce global estimates 
+ Use an online SQL relational database linked to R-scripts

## Talk plan

+ Exploring data for non-random, non-representative sampling (data catalogue)
+ Investigating the relationship (sum of products) between Task 1, which focuses on the total catches of 9 major tuna & tuna-like species (Albacore, Bluefin, Yellowfin, Bigeye, Skipjack, Swordfish, Blue Marlin, White Marlin, Sailfish), and Task 2 data which is more detailed containing more detailed information on location, season, etc.
+ Exploring the relationship between weight and numbers. Note: some countries supply only numbers, some both and some just weights for Task 2
+ Developing multi-variate models (e.g. GAMs) for interpoplation
+ Integrating activities with an online geographic database system (based on Postgis)

## Sampling by year and month - Chinese Taipei

```{r code,include=FALSE}
library(rio)
library(spatial)
library(sp)
library(doBy)
library(rgdal)
library(RODBC)
library(RColorBrewer)
library(ggplot2)
library(vmstools)
library(mgcv)
library(maps)
library(mapdata)
library(reshape2)
library(rgeos)
library(lattice)

source("/home/doug/effdis/R/yr.month.coverage.task2.r")
source("/home/doug/effdis/R/trend.r")
source("/home/doug/effdis/R/spatial.coverage.by.year.task2.r")
t2ce <- read.table("/home/doug/effdis/data/t2ce.csv",sep=",",header=T)
t1det9sp <- import("/home/doug/effdis/data/t1det_9sp.xlsx")
#colnames(t2ce) <- tolower(colnames(t2ce))
#dimnames(t2ce)[[2]][55:56] <- c('longitude','latitude') # Have to be different.
#t2ce$trend <- trend.r(year=t2ce$yearc,month=t2ce$timeperiodid,start.year=1950)

```
```{r temporal confounding Chinese Taipei,include=TRUE,echo=FALSE}
yr.month.coverage.task2.r(tdata=t2ce,which.flag='Chinese Taipei')
```

## Sampling by year and month - Japan

```{r temporal confounding Japan,include=TRUE,echo=FALSE}
yr.month.coverage.task2.r(tdata=t2ce,which.gear='LL',which.flag='Japan')
```

## Sampling in space by year - Chinese Taipei 

```{r spatial confounding China,include=TRUE,echo=FALSE}
library(maps)
par(mfrow=c(5,5),mar=c(0,0,1,0))
spatial.coverage.by.year.task2.r(tdata=t2ce,start.year=1986,end.year=2010,which.gear='LL',which.flag='Chinese Taipei')
```

## No of hooks by year - Chinese Taipei and Japan

```{r no of hooks reported by Taipei and Japan in task 2,include=TRUE,echo=FALSE,message = FALSE, error=FALSE, warnings=FALSE}
source("/home/doug/effdis/R/effort.by.year.task2.r")
par(mfrow=c(2,1),mar=c(3,3,3,3))
effort.by.year.task2.r(which.flag='Chinese Taipei')
effort.by.year.task2.r(which.flag='Japan')
```


## No of hooks by year and location - China P.R.

```{r no of hooks shot by China P.R. in 2006 from task 2,include=TRUE,echo=FALSE,message = FALSE, error=FALSE, warnings=FALSE}
source("/home/doug/effdis/R/three.d.effort.by.year.r")
t2ce_lf_ll_ChinaPR <- read.table('/home/doug/effdis/data/t2ce_lf_ll_ChinaPR.csv',sep=',',header=T)
par(mfrow=c(1,1),mar=c(3,3,3,3))
three.d.effort.by.year.r(tdata=t2ce_lf_ll_ChinaPR,what.year='2006',what.flag='China P.R.',scaling.f=100000)  
```

## Weight of albacore caught by year and location - China P.R.

```{r kg albacore caught by China P.R. in 2006 from task 2,include=TRUE,echo=FALSE,message = FALSE, error=FALSE, warnings=FALSE}
source("/home/doug/effdis/R/three.d.catch.by.year.r")
par(mfrow=c(1,1),mar=c(3,3,3,3))
three.d.catch.by.year.r(tdata=t2ce_lf_ll_ChinaPR,what.year='2006',what.species='alb',what.flag='China P.R.',scaling.f=1000)  
```



## Relationship (sum of products) between Task 1 (the total catches of 9 major tuna & tuna-like species) and Task 2 (much more detailed) data for all fleet combinations


```{r relationship between task 1 and task 2 data, include=TRUE,echo=FALSE,message = FALSE, error=FALSE, warnings=FALSE}
task1.2 <- read.table("/home/doug/effdis/data/task1.2.csv",header=T,sep=',')
xyplot(log(total_catch_kgs) ~ log(measured_catch)|flagname,groups=catchunit,data=task1.2[task1.2$region == 'AT' ,]
       ,xlab='task2',ylab='task1',auto.key = T)
```

## Developing multivariate models. 

```{r regression modeling, include=TRUE,echo=FALSE,message = FALSE, error=FALSE, warnings=FALSE}

mod.output <- read.table('/home/doug/effdis/data/model-data-alb.csv',sep=',',header=T)

lonnie <- seq(-96.5,18.5,by=5)
lattie <- seq(-56.5,58.5,by=5)
lo<- length(lonnie)
la<- length(lattie)
par(mfrow=c(4,4),mar=c(1,1,3,1))

for(i in c(1,3,9,11)){
  
  grd <- mod.output[mod.output$month == i & mod.output$year == 2006,]
  grd$catch <- grd$measured_catch*grd$prob
  
  image(lonnie,lattie,matrix(grd$prob,lo,la),col=topo.colors(100),xlab='',ylab='',xaxt='n',yaxt='n')
  contour(lonnie,lattie,matrix(grd$prob,lo,la),col=topo.colors(100),add=T)
  map('worldHires',add=T,fill=T);title(paste('Probability of catch (P)',month.abb[i],'2006'))
  
  image(lonnie,lattie,matrix(log(grd$measured_catch),lo,la),col=topo.colors(100),xlab='',ylab='',xaxt='n',yaxt='n')
  contour(lonnie,lattie,matrix(log(grd$measured_catch),lo,la),col=topo.colors(100),add=T)
  map('worldHires',add=T,fill=T);title(paste('Catch without zeros (C1)',month.abb[i],'2006'))
  
  image(lonnie,lattie,matrix(log(grd$catch),lo,la),col=topo.colors(100),xlab='',ylab='',xaxt='n',yaxt='n')
  contour(lonnie,lattie,matrix(log(grd$catch),lo,la),col=topo.colors(100),add=T)
  map('worldHires',add=T,fill=T);title(paste('Catch P x C1',month.abb[i],'2006'))
  

  image(lonnie,lattie,matrix(grd$eff,lo,la),col=topo.colors(100),xlab='',ylab='',xaxt='n',yaxt='n')
  contour(lonnie,lattie,matrix(log(grd$eff),lo,la),col=topo.colors(100),add=T)
  map('worldHires',add=T,fill=T);title(paste('No of hooks',month.abb[i],'2006'))
}

```

## First estimate 

```{r global effort estimate, include=TRUE,echo=FALSE,message = FALSE, error=FALSE, warnings=FALSE}

effdis <- read.table('/home/doug/effdis/data/effdis-estimate.csv',sep=',',header=T)
par(mfrow=c(1,1),mar=c(5,5,4,4))
plot(effdis$year,effdis$effort,xlab='',ylab='hooks',type='l',lwd=3)
abline(v=seq(1950,2010,by=10),lty=2,col='blue')
title('Task 1 catch / Task 2 CPUE')
```


## Conclusions
+ The data are complex with changing fleet aggregations etc
+ There is non-random, non-representative sampling
+ There is a strong positive correlation between Task 1 and Task 2 data for most countries which is useful information
+ Provision of numbers for Task 2 (instead of weights) by e.g. USA creates problems
+ Regression modeling is a promising approach
+ A series of models can be fitted to the data and used for interpolation and variance estimation
+ Can effort realistically be estimated without any measure of capacity ?





